Reducing RES Droughts through the integration of wind and PV

### 2025-03-14 Article/RES_droughts_Morin_ReEn_R2.tex
Update Reviewed version of the Article

Atlite and C3S-E datasets are driven by the ERA5 reanalysis~\citep{hersbach2020era5}, produced by the European Centre for Medium-Range Weather Forecasts (ECMWF). This global gridded dataset provides hourly atmospheric variables from 1940 to the present at a horizontal resolution of 0.25\textdegree. It is widely used for estimating solar and wind energy~\citep{mockert2023drought, dubus2023energy, brown2021drought, otero2022drought}. Table~\ref{tab:var_name} lists the ERA5 variables used by Atlite and C3S-Energy.

@@

@@

The seasonal analysis of RES droughts is based on the percentage of hours in each month classified as drought events. For the wind-only scenario (Fig.~\ref{fig:res_droughts_seasonality}a), the Atlite dataset exhibits a pronounced seasonal pattern, with about 24\% of summer hours (June–July–August) identified as droughts compared to only 4\% in winter (December–January–February). This strong seasonal signal is less evident in the C3S-E datasets, which suggests that the differences in the underlying wind power curves play a significant role. In Atlite, CF near or below the 0.1 threshold occur at relatively higher wind speeds, resulting in a higher count of drought hours during the summer months.

>>>>

Atlite and C3S-E datasets are driven by the ERA5 reanalysis~\citep{hersbach2020era5}, produced by the European Centre for Medium-Range Weather Forecasts (ECMWF). This global gridded dataset provides hourly atmospheric variables from 1940 to the present at a horizontal resolution of 0.25\textdegree. It has proven to be the best choice for studying renewable energy in Ireland \citep{doddy2021era5}. Table~\ref{tab:var_name} lists the ERA5 variables used by Atlite and C3S-Energy.

@@

It is important to note that this analysis is not focused on the absolute numbers characterizing RES droughts, but rather on the relative differences observed when comparing the various datasets and energy scenarios described in Section~\ref{sec:scenarios}.

@@

The seasonal analysis of RES droughts is based on the percentage of hours in each month classified as drought events. Wind droughts tend to be more frequent during summer, whereas PV droughts are more common in winter due to reduced sunlight. By comparing these seasonal patterns across different datasets and energy scenarios, the study examines how model-specific assumptions and variations in capacity mix affect the overall characterisation of drought events.

For the wind-only scenario (Fig.~\ref{fig:res_droughts_seasonality}a), the Atlite dataset exhibits a pronounced seasonal pattern, with about 24\% of summer hours (June–July–August) identified as droughts compared to only 4\% in winter (December–January–February). This strong seasonal signal is less evident in the C3S-E datasets, which suggests that the differences in the underlying wind power curves play a significant role. In Atlite, CF near or below the 0.1 threshold occur at relatively higher wind speeds, resulting in a higher count of drought hours during the summer months.
